---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
setwd("E:/Continuing Education/Data Scientist/Reproducible Research")

## Loading and preprocessing the data

Load the packages needed for my version of the assignment.

```{r}
library(ggplot2)
```

Load the data and convert it into a table Data Frame for use with dplyr
```{r}
activityData<-read.csv("activity.csv")
activityData<-tbl_df(activityData)
```

## What is mean total number of steps taken per day?

Subset the data, remove the NA's and group the data by date
```{r}
totalSteps<- tapply(activityData$steps, activityData$date, sum, na.rm = TRUE) 
```

Plot the histogram
```{r}
qplot(totalSteps, xlab = "Total steps", ylab = "Frequency", geom="histogram")
```

Mean number steps per day
```{r}
totalStepsMean<-mean(totalSteps)
```

Median number steps per day
```{r}
totalStepsMedian<-median(totalSteps)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
