---
title: 'Reproducible Research: Peer Assessment 1'
author: "Scott Woosley"
date: "Friday, April 17, 2015"
output: html_document
---

setwd("E:/Continuing Education/Data Scientist/Reproducible Research")

## Loading and preprocessing the data

Load the packages needed for my version of the assignment.

```{r, echo=TRUE}
library(ggplot2)
library(dplyr)
library(lubridate)
```

Load the data and convert it into a table Data Frame for use with dplyr
```{r, echo=TRUE}
activityData<-read.csv("activity.csv")
activityData<-tbl_df(activityData)
```

## What is the mean total number of steps taken per day?

Subset the data, remove the NA's and group the data by date
```{r, echo=TRUE}
totalSteps<- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE) 
```

Plot the histogram
```{r, echo=TRUE}
qplot(totalSteps, xlab = "Total steps", ylab = "Frequency", geom="histogram")
```


```{r, echo=TRUE}
totalStepsMean<-mean(totalSteps)
```
Mean number steps per day: `r totalStepsMean`


```{r, echo=TRUE}
totalStepsMedian<-median(totalSteps)
```
Median number steps per day: `r totalStepsMedian`


## What is the average daily activity pattern?
Make type "1" plot with the 5 minute interval
```{r, echo=TRUE}
averageSteps <- activityData %>%
    filter(interval == 5) %>%
    group_by(date)
plot(averageSteps$steps, averageSteps$date, type="l", xlab = "Average Steps", ylab = "Date")
title(main="Average Steps per Day for the 5 minute Interval", font.main=4)

```

Day 10 has the greatest average number of steps

## Imputing missing values


```{r, echo=TRUE}
rowCount<-nrow(filter(activityData, is.na(steps)))
```
There are `r rowCount` missing values in the dataset

Create an average of daily averages and create a new data set with missing values substituted with the average

```{r, echo=TRUE}
dailyAverage <- activityData %>%
filter(!is.na(steps)) %>%
group_by(date) %>%
summarize(average = mean(steps))
averageAverage <- mean(dailyAverage$average)

activityDataUpdate <- activityData %>%
mutate(steps=ifelse(is.na(steps), averageAverage, steps))
```

Create new Histogram and new average and median results
```{r, echo=TRUE}
totalStepsUpdate<- tapply(activityDataUpdate$steps, activityData$date, sum, na.rm=TRUE)
qplot(totalStepsUpdate, xlab = "Total steps", ylab = "Frequency", geom="histogram")
totalStepsMean <- mean(totalStepsUpdate)
totalStepsMedian <- median(totalStepsUpdate)

```

Mean number steps per day: `r totalStepsMean`

Mean number steps per day: `r totalStepsMedian`


## Are there differences in activity patterns between weekdays and weekends?

```{r}
activityDataWeekday<-mutate(activityDataUpdate, day = wday(ydm(as.character(date))))
activityDataWeekday <- mutate(activityDataWeekday, weekend = ifelse(day == c(1,7), "Weekend", "Weekday"))
activityDataWeekday$weekend <- factor(activityDataWeekday$weekend)

averageStepsWeekday <- activityDataWeekday %>%
    filter(interval == 5, weekend == "Weekday") %>%
    group_by(date)
    
averageStepsWeekend <- activityDataWeekday %>%
    filter(interval == 5, weekend == "Weekend") %>%
    group_by(date)


par(mfrow=c(2,1))
plot(averageStepsWeekday$steps, averageStepsWeekday$day, type="l", xlab = "Average Steps", ylab = "Date")
title(main="Average Steps per Day for the 5 minute Interval on Weekdays", font.main=4)
plot(averageStepsWeekend$steps, averageStepsWeekend$day, type="l", xlab = "Average Steps", ylab = "Date")
title(main="Average Steps per Day for the 5 minute Interval on Weekends", font.main=4)
```


